The function-sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-29 crelle-dictionary.tex
Working through paper, fixing small things

\begin{abstract}
We introduce the category of quasicharacter sheaves on smooth commutative group schemes $G$ over finite fields $k$ and, in doing so, expand the scope of the function-sheaf dictionary from connected commutative algebraic groups to this setting.
We show that the group of isomorphism classes of quasicharacter sheaves is an extension of the group of characters of $G(k)$ by a cohomology group determined by the component group scheme of $G$.
As an application, we find a category of sheaves for which isomorphism classes determine quasi\-characters in the case of algebraic tori, abelian varieties and unipotent wound groups over local fields. 
In this geometrization of quasicharacters, our sheaves are local systems on
Greenberg transforms of locally finite type N\'eron models.
\end{abstract}

@@

We first test the category of quasicharacter sheaves by examining its objects on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale});
\'etale group schemes form a counterpoint to connected algebraic groups, since the component group of any smooth group scheme is an \'etale group scheme.
Our key tools for understanding the trace of Frobenius in the \'etale case are a reinterpretation of $\QC(G)$ in terms of stalks (\S\ref{ssec:stalks}) and a
relationship with the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\QCiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT}) that 
%$\TrFrob{G}$ is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$. We find that 
the group homomorphism 
\[
\TrFrob{G}: \QCiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
One can show, using the K\"unneth formula, that $\Hh^2(\bG,\EEx)$ is non-trivial unless $\bG$ is cyclic.
%The fact that $\TrFrob{G}$ may not be an isomorphism should not be too troubling: if quasicharacter sheaves $(\gqcs{L}, \mu,\phi)$ and $(\gqcs{L}', \mu',\phi')$ determine the same character of $G(\Fq)$ then $(\gqcs{L},\phi)$ and $(\gqcs{L}',\phi')$ are isomorphic as Weil sheaves.

Returning to the case of connected commutative algebraic groups $G$, we consider quasicharacter sheaves that arise via base change to $\bFq$ from local systems on $G$ (\S\ref{ssec:descentG}) and
those that appear in a pushforward from a constant sheaf along a discrete isogeny $H \to G$ (\S\ref{ssec:discrete_isogenies}).  While these constructions make sense even for non-connected $G$, in the connected
case we show that every quasicharacter sheaf can be described in both of these ways (\S\ref{ssec:connected}).  We use this fact to prove that $\TrFrob{G}$ is an isomorphism for connected commutative algebraic groups $G$
and to determine the automorphism groups of quasicharacter sheaves on such $G$.

%
%The appearance of $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ in these two cases is interesting: on the one hand, when $G$ is a connected commutative algebraic group over $\Fq$, isomorphism classes of quasicharacter sheaves on $G$ are measured by $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ and the trace of Frobenius gives an isomorphism $\TrFrob{G}: \QCiso{G} \to \Hom(G(\Fq),\EEx)$; on the other hand, if $G$ an etale commutative group scheme over $\Fq$, then $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ measures the kernel of $\TrFrob{G}: \QCiso{G} \to \Hom(G(\Fq),\EEx)$. 
%This observation makes plain the fact that  . . .
%Indeed, quasicharacter sheaves on \'etale commutative group schemes that are trivialized by discrete isogenies are invisible to the trace of Frobenius.

@@

By any measure, there are more quasicharacter sheaves for $T$ than quasicharacters of $T(K)$.  
In this regard, we are reminded of the work of David Vogan \cite{vogan:93a}, in which he finds a geometrization of complete Langlands parameters for $p$-adic groups 
and, in the process, is led to study the representations of all the pure rational forms of the $p$-adic group, simultaneously.
A similar phenomenon appears in recent work by Joseph Bernstein in which his geometric Ansatz leads to the study of certain sheaves on the stacky classifying space of the $p$-adic group, resulting in a category which appears to be tied to the representations of all the pure rational forms of the $p$-adic group \cite{bernstein:vogan_conference}.
Indeed, Bernstein has suggested to us that our category of quasicharacters for $T$ may be tied to quasicharacters of all the pure rational forms of $T$.

The process of creating a category from the group of quasicharacters of a $p$-adic  torus informs our choice of the term quasicharacter sheaf in this paper.
We also want to situate our terminology in a historical context.
It is not uncommon to refer to local systems satisfying \eqref{introbox} on a connected, commutative algebraic group as character sheaves;
see for example, \cite{kamgarpour:09a}*{Intro}.
As we have indicated, quasicharacter sheaves evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
However, the main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on reductive groups
over algebraically closed fields with finite cyclic component groups in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, such groups are extensions of finite cyclic groups by algebraic tori,
in which case it is not difficult to relate Frobenius-stable character sheaves to our quasicharacter sheaves.
The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0({\bar G}),\EEx)^{\Weil{}}$ do not arise in that context.

@@

We will see that $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ may not provide complete
information about isomorphism classes of quasicharacter sheaves on $G$ when $G$ is not a connected algebraic group.
Our main tool for understanding this phenomenon
is a group homomorphism $S_G: \QCiso{G} \to \Hh^2(E^\bullet_G)$ defined in Section~\ref{ssec:S}, for which the next two sections are preparation.

@@

The zeroth page of the Hochschild-Serre spectral sequence
is a double complex $E^{\bullet, \bullet}$ defined by
\[
E^{i,j} = C^i(\Weil{}, C^j(G(\bFq), \EEx));
\]
see \cite{Vakil:spectral_sequences}, expanding on \cite{Weibel:Homological_Algebra}*{Ch 5 and \S 7.5}.
The standard derivative on cochains yields two derivatives,
\begin{align*}
d_G &: E^{i,j} \to E^{i,j+1} \quad \mbox{and} \\
d_{\Weil{}} &: E^{i,j} \to E^{i+1,j};
\end{align*}
we use the first as the derivative $d_0$ and the second to induce $d_1$.
Combining them also yields a derivative $d = d_G + (-1)^j d_{\Weil{}}$ on the total complex
\[
E^n_G = \bigoplus_{i+j=n} E^{i,j}.
\]
The machinery of spectral sequences gives us a sequence of pages $E_r^{i,j}$, converging to a page $E_{\infty}^{i,j}$. We summarize the key properties of this spectral sequence in the following proposition.

@@

\begin{corollary}\label{cor:spectral_ses}
There is a short exact sequence
 \[
    0 \to
    \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to
    \Hh^2(E^\bullet_G) \to
    \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to
    0.
 \]
\end{corollary}

This sequence will play a key role in understanding the kernel of $\TrFrob{G}$, as described in the next few sections.
For this application, we need a good understanding of these maps to and from the total complex.

\begin{proposition} \label{prop:ses_desc}
Consider the short exact sequence in Corollary~\ref{cor:spectral_ses}.
\begin{enumerate}
\item Every class $[\alpha\oplus\beta\oplus\gamma] \in \Hh^2({\bar G},\EEx)$ is cohomologous to one with $\gamma=0$.
\item The map $\Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx))$ is given by $[\alpha\oplus\beta\oplus 0] \mapsto [\beta]$.
\item Suppose $a \in Z^2({\bar G}, \EEx)$ represents a class in $\Hh^2({\bar G},\EEx)$ fixed by Frobenius.
The map $\Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G)$ is given by $[a] \mapsto [a \oplus 0 \oplus 0]$.
\end{enumerate}
\end{proposition}
\begin{proof}
Since $\Hh^2(\Weil{}, C^0({\bar G}, \EEx)) = 0$, we may find a $\gamma_1 \in C^1(\Weil{}, C^0({\bar G}, \EEx))$ with $d_{\Weil{}}\gamma_1 = \gamma$.
Subtracting $d \gamma_1$ from $\alpha\oplus\beta\oplus\gamma$, we may assume that $\gamma = 0$.

@@

In this section we define an isomorphism
\[
S_G : \QCiso{G} \to \Hh^2(E^\bullet_G).
\]
Let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
For each geometric point $x\in {\bar G}$, choose a basis $\{ v_x \}$ for $\gqcs{L}_x$.
Through this choice, $\qcs{L}$ determines functions
\begin{align*}
a : {\bar G}\times {\bar G} &\to \EEx & b : {\bar G} &\to \EEx \\
\mu_{x,y}(v_{x+y}) &= a(x,y) v_x \otimes v_y & \phi_x(v_{\Frob{G}(x)}) &= b(x) v_x.
\end{align*}
Condition~\ref{QC.2} implies that
\begin{equation}\label{2-cocyle}
a(x+y,z) a(x,y) = a(x,y+z) a(y,z)
\end{equation}
for all $x,y,z\in {\bar G}$, so $a \in Z^2({\bar G},\EEx)$.  Similarly, condition~\ref{QC.3} gives
\begin{equation}\label{nohom}
\frac{a(\Frob{G}(x),\Frob{G}(y))}{a(x,y)} =  \frac{b(x+y)}{b(x) b(y)}
\end{equation}
for $x, y \in {\bar G}$.
Let $\alpha \in C^0(\Weil{},C^2({\bar G},\EEx)$ be the $0$-cochain corresponding to $a$ and let $\beta\in C^1(\Weil{},C^1({\bar G},\EEx)$ be the cocycle such that $\beta(\Frob{})$ is $b$.  We will write both $\alpha$ and $\beta$ additively.
Then
\[
d_G\alpha =0, \qquad\qquad
d_{\Weil{}} \alpha = d_{G} \beta,\qquad\qquad
d_{\Weil{}} \beta =0;
\]
in other words,
\[\alpha\oplus \beta \in Z^2(E^\bullet_G).\]
Although the cocycle $\alpha\oplus \beta$ is not well defined by $\qcs{L}$, its class in $Z^2(E^\bullet_G)$ is.
To see this, let $\{ v'_x \in \gqcs{L}_x^\times \tq x \in {\bar G}\}$ be another choice and let $\alpha'\oplus \beta' \in Z^2(E^\bullet_G)$ be defined by $\qcs{L}$ and this choice, as above.
Now let $\delta \in C^0(\Weil{},C^1({\bar G},\EEx))$ correspond to the function $d : {\bar G}\to \EEx$ defined by $v'_x = d(x) v_x$.
Chasing through \ref{QC.2} and \ref{QC.3}, we find
\[
\alpha'\oplus\beta' = \alpha\oplus\beta + d\delta,
\]
so the class $[\alpha\oplus\beta]$ of $\alpha\oplus\beta$ in $\Hh^2(E^\bullet_G)$ is independent of the choice made above. It is also easy to see that $[\alpha\oplus\beta] = [\alpha_0\oplus\beta_0]$ when $\qcs{L} \iso \qcs{L}_0$,
which concludes the definition of the function
\begin{align*}
S_G : \QCiso{G} &\to \Hh^2(E^\bullet_G)\\
[\qcs{L}] &\mapsto [\alpha\oplus \beta].
\end{align*}
It is also easy to see that $[\alpha_1\oplus\beta_1] + [\alpha_2\oplus\beta_2] = [\alpha_3\oplus\beta_3]$ when $\qcs{L}_3 = \qcs{L}_1\otimes \qcs{L}_2$, so $S_G$ is a group homomorphism.

@@

\begin{proposition}\label{prop:SGiso}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $[\qcs{L}] \in \QCiso{G}$ with $S_G([\qcs{L}]) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EE$ by $\sigma_x : v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{proposition}\label{prop:sur_etale}
If $G$ is \'etale, then $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is surjective
and admits a section.
\end{proposition}
\begin{proof}
Pick $\chi \in G(\Fq)^*$. 
Let $[\beta]\in \Hh^1(\Weil{},\bG^*)$ be the class corresponding to $\chi$ under Lemma~\ref{lem:dual-inv}.
Every representative cocycle $\beta \in Z^1(\Weil{},\bG^*)$ determines a homomorphism $\beta(\Frob{}) : G(\bFq)\to \EEx$ such that $\beta(\Frob{})\vert_{G(\Fq)} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$ and
$\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = \beta(\Frob{})(x)$.
Since $\beta(\Frob{}) : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $a =1$.
So $\qcs{L} = (\gqcs{L}, \mu, \phi)$
is a quasicharacter sheaf with $\trFrob{\qcs{L}} = \chi$.
This shows that $\TrFrob{G}$ is surjective.
%
Now let $\beta' \in Z^1(\Weil{},\bG^*)$ be another representative for $[\beta]$
so $\beta'-\beta = d_{\Weil{}} \delta$ for some $\delta \in C^0(\Weil{},\bG^*)$ defining $d \in \Hom(G(\bFq),\EEx)$.
Let $\qcs{L}'$ be the quasicharacter sheaf on $G$ defined by $\beta'$, as above.
For each $x\in G(\bFq)$, define $\brho_x :\qcs{L}_x\to \qcs{L}'_x$ by $\brho_x(1) = d(x)$.
Since condition~\ref{qc.4} is satisfied for $\{ \brho_x \tq x\in G(\bFq)\}$, it defines a morphism $\rho : \qcs{L}\to \qcs{L}'$, which is clearly and isomorphism. 
%
We have now defined a section of $\TrFrob{G}$. 
\end{proof}

\begin{proposition} \label{prop:etale-iso}
 If $G$ is \'etale then the map $S_G : \QCiso{G}\to \Hh^2(E^\bullet_G)$ induces an isomorphism of short exact sequences
\[
\begin{tikzcd}
 0 \arrow{r} & \ker \TrFrob{G} \arrow{d} \arrow{r} & \QCiso{G}\arrow{d}{S_G} \arrow{r}{\TrFrob{G}} \arrow{r} & G(\Fq)^* \arrow{d} \arrow{r} & 0\\
  0 \arrow{r} & \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \arrow{r} & \Hh^2(E^\bullet_G) \arrow{r} & \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \arrow{r} & 0.
 \end{tikzcd}
 \]
\end{proposition}
\begin{proof}
This result follows easily from Propositions~\ref{prop:ses_desc}, \ref{prop:SGiso} and \ref{prop:sur_etale}.
\end{proof}

Combining Propositions~\ref{prop:sur_etale} and \ref{prop:etale-iso} gives the following

\begin{corollary}\label{cor:split}
The short exact sequence
 \[
    0 \to
    \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to
    \Hh^2(E^\bullet_G) \to
    \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to
    0
 \]
 is split.
\end{corollary}

@@

Most of this paper focuses on the question of determining the isomorphism classes of objects; a complete understanding of the morphisms in $\QC(G)$ also requires a description of the automorphisms of an arbitrary quasicharacter sheaf $\qcs{L}$.

@@

Conversely, if $\rho : G(\bFq)_{\Weil{}} \to \EEx$ is any homomorphism then defining $\brho_x$ as multiplication by $\rho(x)$ will define $\gqcs{L} \to \gqcs{L}'$ satisfy the two diagrams in \ref{qc.4}.

@@

\begin{definition}
Let $\bQC(G)$ be the category of pairs $(\qcs{E},\mu_\qcs{E})$
where $\qcs{E}$ an $\ell$-adic local system on $G$ of rank-one,
equipped with an isomorphism $\mu_\qcs{E} : m^* \qcs{E} \to \qcs{E} \boxtimes \qcs{E}$
satisfying the analogue of \ref{QC.1} on $G$;
morphisms in $\bQC(G)$ are defined as in the second part of
\ref{QC.3}.
\end{definition}

>>>>

\begin{abstract}
We introduce the category of quasicharacter sheaves on smooth commutative group schemes $G$ over finite fields $k$ and expand the scope of the function-sheaf dictionary from connected commutative algebraic groups to this setting.
We show that the group of isomorphism classes of quasicharacter sheaves is an extension of the group of characters of $G(k)$ by a cohomology group determined by the component group scheme of $G$.
As an application, we find a category of sheaves whose isomorphism classes correspond to quasicharacters in the case of algebraic tori, abelian varieties and unipotent wound groups over local fields. 
In this geometrization of quasicharacters, our sheaves are local systems on
Greenberg transforms of locally finite type N\'eron models.
\end{abstract}

@@

We first test the category of quasicharacter sheaves by examining its objects on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale});
\'etale group schemes form a counterpoint to connected algebraic groups, since the component group of any smooth group scheme is an \'etale group scheme.
Our key tools for understanding the trace of Frobenius in the \'etale case are a reinterpretation of $\QC(G)$ in terms of stalks (\S\ref{ssec:stalks}) and a
relationship with the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\QCiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT}) that 
%$\TrFrob{G}$ is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$. We find that 
the group homomorphism 
\[
T_G: \QCiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
One can show, using the K\"unneth formula, that $\Hh^2(\bG,\EEx)$ is non-trivial unless $\bG$ is cyclic.
The fact that $T_G$ may not be an isomorphism should not be too troubling: if quasicharacter sheaves $(\gqcs{L}, \mu,\phi)$ and $(\gqcs{L}', \mu',\phi')$
determine the same character of $G(\Fq)$ then $(\gqcs{L},\phi)$ and $(\gqcs{L}',\phi')$ are isomorphic as Weil sheaves.

Returning to the case of connected commutative algebraic groups $G$, we consider quasicharacter sheaves that arise via base change to $\bFq$ from local systems on $G$ (\S\ref{ssec:descentG}) and
those that appear in a pushforward from a constant sheaf along a discrete isogeny $H \to G$ (\S\ref{ssec:discrete_isogenies}).  While these constructions make sense even for non-connected $G$, in the connected
case we show that every quasicharacter sheaf can be described in both of these ways (\S\ref{ssec:connected}).  We use this fact to prove that $T_G$ is an isomorphism for connected commutative algebraic groups $G$
and to determine the automorphism groups of quasicharacter sheaves on such $G$.

%
%The appearance of $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ in these two cases is interesting: on the one hand, when $G$ is a connected commutative algebraic group over $\Fq$, isomorphism classes of quasicharacter sheaves on $G$ are measured by $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ and the trace of Frobenius gives an isomorphism $T_G: \QCiso{G} \to \Hom(G(\Fq),\EEx)$; on the other hand, if $G$ an etale commutative group scheme over $\Fq$, then $\Hh^2(\bG,\EEx)^{\Gal(\bFq/\Fq)}$ measures the kernel of $T_G: \QCiso{G} \to \Hom(G(\Fq),\EEx)$. 
%This observation makes plain the fact that  . . .
%Indeed, quasicharacter sheaves on \'etale commutative group schemes that are trivialized by discrete isogenies are invisible to the trace of Frobenius.

@@

By any measure, there are more quasicharacter sheaves for $T$ than quasicharacters of $T(K)$.  
In this regard, we are reminded of the work of David Vogan \cite{vogan:93a}, in which he finds a geometrization of complete Langlands parameters for $p$-adic groups 
and, in the process, is led to study the representations of all the pure rational forms of the $p$-adic group simultaneously.
A similar phenomenon appears in recent work by Joseph Bernstein in which his geometric Ansatz leads to the study of certain sheaves on the stacky classifying space of the $p$-adic group, resulting in a category which appears to be tied to the representations of all the pure rational forms of the $p$-adic group \cite{bernstein:vogan_conference}.
Indeed, Bernstein has suggested to us that our category of quasicharacters for $T$ may be tied to quasicharacters of all the pure rational forms of $T$.

The process of creating a category from the group of quasicharacters of a $p$-adic  torus informs our choice of the term quasicharacter sheaf in this paper.
We also want to situate our terminology in a historical context.
It is not uncommon to refer to local systems satisfying \eqref{introbox} on a connected, commutative algebraic group as character sheaves;
see for example, \cite{kamgarpour:09a}*{Intro}.
As we have indicated, quasicharacter sheaves evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
However, the main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on reductive groups
over algebraically closed fields with finite cyclic component groups in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, such groups are extensions of finite cyclic groups by algebraic tori,
in which case it is not difficult to relate Frobenius-stable character sheaves to our quasicharacter sheaves.
The new features that we have found pertaining to Weil sheaves and the non-triviality of $\Hh^2(\pi_0({\bar G}),\EEx)^{\Weil{}}$ do not arise in this context.

@@

We will see that $T_G : \QCiso{G} \to G(\Fq)^*$ may not provide complete
information about isomorphism classes of quasicharacter sheaves on $G$ when $G$ is not a connected algebraic group.
Our main tool for understanding this phenomenon
is a group homomorphism $S_G: \QCiso{G} \to \Hh^2(E^\bullet_G)$ defined in Section~\ref{ssec:S}, for which the next two sections are preparation.

@@

The zeroth page of the Hochschild-Serre spectral sequence
is a double complex $E^{\bullet, \bullet}$ defined by
\[
E^{i,j} = C^i(\Weil{}, C^j(G(\bFq), \EEx));
\]
see \cite{Vakil:spectral_sequences}, expanding on \cite{Weibel:Homological_Algebra}*{Ch 5 and \S 7.5}.
The standard derivative on cochains yields two derivatives,
\begin{align*}
d_G &: E^{i,j} \to E^{i,j+1} \quad \mbox{and} \\
d_{\Weil{}} &: E^{i,j} \to E^{i+1,j};
\end{align*}
we use the first as the derivative $d_0$ on the zeroth page, and the second to induce $d_1$.
Combining them also yields a derivative $d = d_G + (-1)^j d_{\Weil{}}$ on the total complex
\[
E^n_G = \bigoplus_{i+j=n} E^{i,j}.
\]
The machinery of spectral sequences gives us a sequence of pages $E_r^{i,j}$, converging to a page $E_{\infty}^{i,j}$. We summarize the key properties of this spectral sequence in the following proposition.

@@

\begin{corollary}
There is a short exact sequence
 \[
    0 \to
    \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to
    \Hh^2(E^\bullet_G) \to
    \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to
    0
 \]
\end{corollary}

This sequence will play a key role in understanding the kernel of $T_G$, as described in the next few sections.
For this application, we need a good understanding of these maps to and from the total complex.

\begin{proposition} \label{prop:ses_desc}
$ $
\begin{enumerate}
\item Every class $[\alpha\oplus\beta\oplus\gamma] \in \Hh^2(E^\bullet_G)$ is cohomologous to one with $\gamma=0$.
\item The map $\Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx))$ is given by $[\alpha\oplus\beta\oplus 0] \mapsto [\beta]$.
\item Suppose $a \in Z^2({\bar G}, \EEx)$ represents a class in $\Hh^2({\bar G},\EEx)$ fixed by Frobenius.
The map $\Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G)$ is given by $[a] \mapsto [a \oplus 0 \oplus 0]$.
\end{enumerate}
\todo{add more connecting prose}
\end{proposition}
\begin{proof}
Since $\Hh^2(\Weil{}, C^0({\bar G}, \EEx)) = 0$, we may find a $\gamma_1 \in C^1(\Weil{}, C^0({\bar G}, \EEx))$ with $d_{\Weil{}}\gamma_1 = \gamma$.
Subtracting $d \gamma_1$ from $\alpha\oplus\beta\oplus\gamma$, we may assume that $\gamma = 0$.

@@

In this section we define an isomorphism
\[
S_G : \QCiso{G} \to \Hh^2(E^\bullet_G).
\]
Let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
For each geometric point $x\in {\bar G}$, choose a basis $\{ v_x \}$ for $\gqcs{L}_x$.
Through this choice, $\qcs{L}$ determines functions
\begin{align*}
a : {\bar G}\times {\bar G} &\to \EEx & b : {\bar G} &\to \EEx \\
\mu_{x,y}(v_{x+y}) &= a(x,y) v_x \otimes v_y & \phi_x(v_{\Frob{G}(x)}) &= b(x) v_x.
\end{align*}
Condition~\ref{QC.2} implies that
\begin{equation}\label{2-cocyle}
a(x+y,z) a(x,y) = a(x,y+z) a(y,z)
\end{equation}
for all $x,y,z\in {\bar G}$, so $a \in Z^2({\bar G},\EEx)$.  Similarly, condition~\ref{QC.3} gives
\begin{equation}\label{nohom}
\frac{a(\Frob{G}(x),\Frob{G}(y))}{a(x,y)} =  \frac{b(x+y)}{b(x) b(y)}
\end{equation}
for $x, y \in {\bar G}$.
Let $\alpha \in C^0(\Weil{},C^2({\bar G},\EEx)$ be the $0$-cochain corresponding to $a$ and let $\beta\in C^1(\Weil{},C^1({\bar G},\EEx)$ be the cocycle such that $\beta(\Frob{})$ is $b$.  We will write both $\alpha$ and $\beta$ additively.
Then
\[
d_G\alpha =0, \qquad\qquad
d_{\Weil{}} \alpha = d_{G} \beta,\qquad\qquad
d_{\Weil{}} \beta =0;
\]
in other words,
\[\alpha\oplus \beta \in Z^2(E^\bullet_G).\]
Although the cocycle $\alpha\oplus \beta$ is not well defined by $\qcs{L}$, its class in $\Hh^2(E^\bullet_G)$ is.
To see this, let $\{ v'_x \in \gqcs{L}_x^\times \tq x \in {\bar G}\}$ be another choice and let $\alpha'\oplus \beta' \in Z^2(E^\bullet_G)$ be defined by $\qcs{L}$ and this choice, as above.
Now let $\delta \in C^0(\Weil{},C^1({\bar G},\EEx))$ correspond to the function $d : {\bar G}\to \EEx$ defined by $v'_x = d(x) v_x$.
Chasing through \ref{QC.2} and \ref{QC.3}, we find
\[
\alpha'\oplus\beta' = \alpha\oplus\beta + d\delta,
\]
so the class $[\alpha\oplus\beta]$ of $\alpha\oplus\beta$ in $\Hh^2(E^\bullet_G)$ is independent of the choice made above. It is also easy to see that $[\alpha\oplus\beta] = [\alpha_0\oplus\beta_0]$ when $\qcs{L} \iso \qcs{L}_0$,
which concludes the definition of the function
\begin{align*}
S_G : \QCiso{G} &\to \Hh^2(E^\bullet_G)\\
[\qcs{L}] &\mapsto [\alpha\oplus \beta].
\end{align*}
It is also easy to see that $[\alpha_1\oplus\beta_1] + [\alpha_2\oplus\beta_2] = [\alpha_3\oplus\beta_3]$ when $\qcs{L}_3 = \qcs{L}_1\otimes \qcs{L}_2$, so $S_G$ is a group homomorphism.

@@

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $[\qcs{L}] \in \QCiso{G}$ with $S_G([\qcs{L}]) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EE$ by $\sigma_x : v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{proposition}\label{prop:sur_etale}
If $G$ is \'etale, then $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is surjective.
\end{proposition}
\begin{proof}
Pick $\chi \in G(\Fq)^*$.
By Lemma~\ref{lem:dual-inv}, there is some $b\in G(\bFq)^*$ such that $b\vert_{G(\Fq)} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$ and
$\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = b(x)$.
Since $b : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $a =1$ and $\qcs{L} = (\gqcs{L}, \mu, \phi)$
is a quasicharacter sheaf with $\trFrob{\qcs{L}} = \chi$.
\end{proof}

\begin{proposition} \label{prop:etale-iso}
 If $G$ is \'etale then the map $S_G : \QCiso{G}\to \Hh^2(E^\bullet_G)$ induces an isomorphism of short exact sequences
\[
\begin{tikzcd}
 0 \arrow{r} & \ker \TrFrob{G} \arrow{d} \arrow{r} & \QCiso{G}\arrow{d}{S_G} \arrow{r}{\TrFrob{G}} \arrow{r} & G(\Fq)^* \arrow{d} \arrow{r} & 0\\
  0 \arrow{r} & \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \arrow{r} & \Hh^2(E^\bullet_G) \arrow{r} & \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \arrow{r} & 0.
 \end{tikzcd}
 \]
\end{proposition}
\begin{proof}
This result follows from Propositions~\ref{prop:ses_desc} and \ref{prop:sur_etale}.
\end{proof}

@@

Most of this paper focuses on the question of determining the isomorphism classes of objects in $\QC{G}$.
A full understanding of the category also requires a description of the morphisms.

@@

Conversely, if $\rho : G(\bFq)_{\Weil{}} \to \EEx$ is any homomorphism, then defining $\brho_x$ as multiplication by $\rho(x)$ will define a morphism $\gqcs{L} \to \gqcs{L}'$ satisfy the two diagrams in \ref{qc.4}.

@@

\begin{definition}
Let $\bQC(G)$ be the category of pairs $(\qcs{E},\mu_\qcs{E})$
where $\qcs{E}$ is an $\ell$-adic local system on $G$ of rank-one,
equipped with an isomorphism $\mu_\qcs{E} : m^* \qcs{E} \to \qcs{E} \boxtimes \qcs{E}$
satisfying the analogue of \ref{QC.2} on $G$;
morphisms in $\bQC(G)$ are defined as in the second part of
\ref{QC.4}.
\end{definition}
